Generative Models for scRNA

### 2018-03-23 generativemodelsforscrna/outline.tex
Added plate diagram

@@

\begin{enumerate}
\item $T^c \sim \Categorical(\pi)$, one of the leaves of $H$.  Let $t_0,t_1(T^c),t_2(T^c),\cdots ,t_\ell^c(T^c)$, indicate the nodes in the unique path in $H$ which leads from the root to the leaf $T$.  Here $t_0$ is the root of the tree, $\ell^c$ is the depth of the leaf $T^c$, and $t^c_\ell(T^c)=T^c$.   To consolidate notation, we take $T^c_i \triangleq t_i(T^c)$.
\item For each $i<\ell$ let $Z^c_{T^c_i} \sim \cN(\mu_{T^c_{i+1}},\Sigma_{T^c_{i+1}})$, and take $Z^c_{T^c_\ell} \sim \cN(0,I)$
\item For each gene $g$, let $\lambda^c_g \in \mathbb R^{p}$ denote a set of parameters.  These will be a deterministic function of the $Z$s:
\[
\lambda^c_g = f_g \left(\sum_{i=0}^{\ell} h_{T^c_i}(Z^c_{T^c_i})\right)
\]
\item Conditioned on $Z^c$ we will have $X^c_g \sim p(x_g ; \lambda^c_g)$. 
\end{enumerate}

Let us unpack some of this:

\begin{itemize}
\item The latent representation $Z^c_i$ corresponds to features of that cell which are specific to all the cells which are descendents of $T^c_i$.
\item The the cell's subcluster $T^c_{i+1}$ effects the distribution of its latent features at the $i$th level, $Z^c_i$.  That is, the latent distribution on the $i$th latent representation is governed by what subcluster it is part of within $T^c_i$.
\item The distribution of the final latent representation, $Z^c_T$, is simply a standard normal.
\item The functions $h_{T^c_i}$ indicates how these cell-type-specific features alter the distribution on the gene expressions for cells in that branch of the tree.  
\end{itemize}

>>>>

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/plate}
\caption{Plate diagram for the model \label{fig:plate}}
\end{figure}

@@

\begin{enumerate}
\item $T^c \sim \Categorical(\pi)$, one of the leaves of $H$.  Let $t_0,t_1(T^c),t_2(T^c),\cdots ,t_\ell^c(T^c)$, indicate the nodes in the unique path in $H$ which leads from the root to the leaf $T$.  Here $t_0$ is the root of the tree, $\ell^c$ is the depth of the leaf $T^c$, and $t^c_\ell(T^c)=T^c$.   To consolidate notation, we take $T^c_i \triangleq t_i(T^c)$.
\item For each $i<\ell$ let $Z^c_{T^c_i} \sim \cN(\mu_{T^c_{i+1}},\Sigma_{T^c_{i+1}})$, and take $Z^c_{T^c_\ell} \sim \cN(0,I)$
\item For each gene $g$, let $\lambda^c_g \in \mathbb R^{p}$ denote a set of parameters.  These will be a deterministic function of the $Z$s:
\[
\lambda^c_g = f_g \left(\sum_{i=0}^{\ell} h_{g,T^c_i}(Z^c_{T^c_i})\right)
\]
\item Conditioned on $Z^c$ we will have $X^c_g \sim p(x_g ; \lambda^c_g)$. 
\end{enumerate}

This can be a bit confusing to take in.  The plate diagram in Figure \ref{fig:plate} may help.  We can also try to unpack some of this in words:

\begin{itemize}
\item The latent representation $Z^c_i$ corresponds to features of that cell which are specific to all the cells which are descendents of $T^c_i$.
\item The the cell's subcluster $T^c_{i+1}$ effects the distribution of its latent features at the $i$th level, $Z^c_i$.  That is, the latent distribution on the $i$th latent representation is governed by what subcluster it is part of within $T^c_i$.
\item The distribution of the final latent representation, $Z^c_T$, is simply a standard normal.
\item The functions $h_{g,T^c_i}$ indicates how these cell-type-specific features alter the distribution on the gene expressions for cells in that branch of the tree.  
\end{itemize}
